---
title: "008_Hierarchical_clustering"
author: "yao"
date: "4/12/2022"
output: html_document
---
Load `twice` and other libraries. You need to load `plyr` first and then 
`dplyr` to avoid conflict.

```{r}
library(twice)
library(plyr)
library(dplyr)
```

The first part, we visualize the differentially expressed krab-znfs comparing
in three species. Load gene expression data and human krab-znfs reference:
```{r}
hmchimp <- read.csv("~/github/kap1/hmchimp_results/geneDESeq2Log2.csv")
hmoran <- read.csv("~/github/kap1/hmoran_results/geneDESeq2Log2.csv")
data("hmKZNFs337")

```

Filter gene expression to keep only KZNFs and merge by rownames to keep all.
```{r}
hmchimp <- hmchimp[rownames(hmchimp)%in%hmKZNFs337$ensembl_gene_id,]
hmoran <- hmoran[rownames(hmoran) %in% hmKZNFs337$ensembl_gene_id,]

repeatList <- intersect(rownames(hmchimp), rownames(hmoran))
hmoran <- hmoran[!rownames(hmoran) %in% repeatList, ]

```

Bind all three species sample together
```{r}
df <- rbind.fill(hmchimp, hmoran)

rownames(df) <- c(rownames(hmchimp), rownames(hmoran))
colnames(df) <- c("H1", "H2", "H3", "C1", "C2", "C3", "O1", "O2", "O3")

```

Normalized in log2 scale and visualize the heatmap.
```{r}
log2df <- log2(df)

#plot result
kznfsComplexHeatmap(log2df)
```


Second part is to visualize transposable elements
Load differentially expressed TEs files and select DE TEs in a list `te_list`
```{r}
# select de genes in human vs. chimpanzee
hmchimpTEres <- read.csv("~/github/kap1/hmchimp_results/teDESeq2results.csv")

hmchimpTEres_de <- hmchimpTEres %>% 
    filter(padj < 0.05) %>%
    filter(abs(log2FoldChange) > 1.2)
hmchimpTEres_de_list <- rownames(hmchimpTEres_de)

# select de genes in human vs. orangutans
hmoranTEres <- read.csv("~/github/kap1/hmoran_results/teDESeq2results.csv")
hmoranTEres_de <- hmoranTEres %>%
    filter(padj < 0.05) %>%
    filter(abs(log2FoldChange) > 1.2)
hmoranTEres_de_list <- rownames(hmoranTEres_de)

# combined data
te_list <- intersect(hmchimpTEres_de_list, hmoranTEres_de_list)
```

Load expression data and visualize the top differentially 50 TEs based on variances 
```{r}
hmchimpTE <- read.csv("~/github/kap1/hmchimp_results/teDESeq2Log2.csv")
hmoranTE <- read.csv("~/github/kap1/hmoran_results/teDESeq2Log2.csv")

dfTE <- merge(hmchimpTE, hmoranTE[,c(4,5,6)], by = "row.names", all = FALSE)
rownames(dfTE) <- dfTE$Row.names
dfTE <- dfTE[,-1]

dfTE <- dfTE[rownames(dfTE) %in% te_list,]
colnames(dfTE) <- c("H1", "H2", "H3", "C1", "C2", "C3", "O1", "O2", "O3")

teComplexHeatmap(as.matrix(log2(dfTE)), fileName = "~/TotalbCellsDETEHeatmap", top=50)

```

Next, we visualize and classified TEs based on class levels to have a better 
resolution in each family.
```{r}
data(hg19rmsk_info)

hg19rmsk_de <- hg19rmsk_info[hg19rmsk_info$gene_id %in% te_list, ]

```

Create transposable elements class lists to filter data. Here we separate to 
LTR and non-LTR because LTR has 56 rows and the others are total 39 rows
```{r}
hg19_LTR_de <- hg19rmsk_de[hg19rmsk_de$class_id == "LTR", ] #56
hg19_non_LTR_de <- hg19rmsk_de[hg19rmsk_de$class_id != "LTR", ] #39

```


Normalized expression data with log2(x+1), and visualize differentailly 
expressed LTR, non-LTR TEs.
```{r}
dfTElog2 <- as.matrix(log2(dfTE + 1))

teComplexHeatmap(
    dfTElog2,
    teFilter = hg19_non_LTR_de$gene_id,
    teAnnot = hg19_non_LTR_de,
    fileName = "img/deTEsnonLTR.pdf"
)
```
```{r}
teComplexHeatmap(
    dfTElog2,
    teFilter = hg19_LTR_de$gene_id,
    top = 56,
    teAnnot = hg19_LTR_de[,c(1,2)],
    fileName = "img/deTEsLTR.pdf"
)
```

