# Subset of coordinates

In this file, I will select the coordinates and send them to Ana to do the analysis again.
(1) Subset only the information from hg19 ensGene and rmsk file.
(2) Subsets the name of gene and TEs from TEKRABber result
(3) Generate 6 tables for Ana (human (gene, te), chimpanzee (gene, te), oran (gene, te))

```{r load reference file, eval=FALSE}
library(dplyr)
# hg19 gene reference
hg19gtf <- rtracklayer::import('../hg19.ensGene.gtf')
hg19ens <- as.data.frame(hg19gtf)

# hg19 TE reference
hg19rmskgtf <- rtracklayer::import('../hg19_rmsk_TE.gtf')
hg19rmsk <- Repitools::annoGR2DF(hg19rmskgtf)

# rearrange the data 
hg19ensTable <- hg19ens[,c("gene_id", "seqnames", "start", "end")]
hg19rmskTable <- hg19rmsk[, c("gene_id", "chr", "start", "end")]

```

```{r convert ID, eval=FALSE}
# convert id to gene name 
library(gprofiler2)
geneName <- gconvert(query = unique(hg19ensTable$gene_id), 
                     organism = "hsapiens",target="ENSG",
                     mthreshold = Inf, filter_na = TRUE)

hg19ensTableMerge <- merge(geneName[,c("input", "name")], hg19ensTable,
                           by.x = "input", by.y="gene_id")

```

Then we filter our data, but we write a function and load the data first.
```{r, eval=FALSE}
sigData <- function(df, refTable, filterTE=FALSE) {
    
    df <- df %>% 
            filter(pvalue < 0.05)
    
    df.output <- data.frame()
    
    if (filterTE==FALSE) {
        df.output <- refTable[refTable$name %in% df$geneName, ]
    } else if (filterTE==TRUE) {
        df.output <- refTable[refTable$gene_id %in% df$teName, ]
    }  
    
    df.output <- df.output[!duplicated(df.output), ]
    df.output
}

hmchimpCorr <- read.csv("hmchimp_results/hmCorrResult.csv")
chimpCorr <- read.csv("hmchimp_results/chimpCorrResult.csv")
hmoranCorr <- read.csv("hmoran_results/hmCorrResult.csv")
oranCorr <- read.csv("hmoran_results/hmCorrResult.csv")
```


```{r filter human data, eval=FALSE}
hmchimpGene <- sigData(hmchimpCorr, hg19ensTableMerge, filterTE = FALSE)
hmchimpTE <- sigData(hmchimpCorr, hg19rmskTable, filterTE = TRUE)

hmoranGene <- sigData(hmoranCorr, hg19ensTableMerge, filterTE = FALSE)
hmoranTE <- sigData(hmoranCorr, hg19rmskTable, filterTE = TRUE)

hmGeneMerge <- full_join(hmchimpGene, hmoranGene)
hmTEMerge <- full_join(hmchimpTE, hmoranTE)
```
```{r filter chimpanzee and orangutan data}
#chimpanzee
chimpGene <- sigData(chimpCorr, hg19ensTableMerge, filterTE = FALSE)
chimpTE <- sigData(chimpCorr, hg19rmskTable, filterTE = TRUE)

#orangutan
oranGene <- sigData(oranCorr, hg19ensTableMerge, filterTE = FALSE)
oranTE <- sigData(oranCorr, hg19rmskTable, filterTE = TRUE)
```

Finally, save the six tables
```{r, eval=FALSE}
write.table(hmGeneMerge, file = "coordinates/hmKZNFsCorrd.csv", sep=",")
write.table(hmTEMerge, file = "coordinates/hmTEsCoord.csv", sep=",")
write.table(chimpGene, file = "coordinates/chimpKZNFsCorrd.csv", sep=",")
write.table(chimpTE, file = "coordinates/chimpTEsCorrd.csv", sep=",")
write.table(oranGene, file="coordinates/oranKZNFsCoord.csv", sep=",")
write.table(oranTE, file="coordinates/oranTEsCoord.csv", sep=",")
```

