# correlation percentage

Here we try to shuffle data to see the gene have negative or positive 
correlations comparing to approx. 50% KRAB-ZNFs (42 out of 81) in human. 
Design:  
1. remove krab-znfs from data
2. 

Load pacakge
```{r}
library(TEKRABber)
library(twice)
library(dplyr)
```

Load data and transform to the format for `corrOrthologTE()`

```{r}
data("hmKZNFs337")
gene <- read.csv("counts/hm_gene.csv")
te <- read.csv("counts/hm_te.csv")

# remove krab-znfs from gene
gene_filter <- gene[!(gene$gene %in% hmKZNFs337$ensembl_gene_id), ]
rownames(gene_filter) <- gene_filter$gene
gene_filter <- gene_filter[,c(2,3,4)]

# set row names of TE
rownames(te) <- te$name
te <- te[,c(4,5,6)]

```

Here we set `set.seed(84)` for reproducibility, and calculate correlation.

```{r}
set.seed(84)

iter <- c()
percentage <- c()

for (i in 1:3){
    idx <- sample(8838, size=80)
    gene_table <- gene_filter[idx, ]
    #myfile <- paste0(format(Sys.time(), "%s_%b_%d", ".csv"))
    
    result <- corrOrthologTE(
        gene_table,
        te
    )
    
    count <- length(unique(result$geneName))
    result.sig <- result %>%
        filter(padj < 0.05)
    count.sig <- length(unique(result.sig$geneName))
    iter <- c(iter, as.character(i))
    percentage <- c(percentage, count.sig/count)
}

df_output <- data.frame(
    num = iter,
    per = percentage
)

```

#### Analyze 200 time of percentage
Here we take `set.seed(84)` first 200 result of percentage
```{r}
df200 <- read.csv("hm_results/percentage_result/percentage.csv")
head(df200)
```

# plot result
```{r}
# plot histogram
library(tidyverse)
library(hrbrthemes)

p <- df200 %>%
    ggplot(aes(x=per)) +
    geom_histogram(binwidth=0.02, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    geom_density(color="#e1744d", alpha=0.8) +
    ggtitle("Siginificantly Correlation (200 iterations)") +
    geom_vline(xintercept = 0.519, linetype="dashed", color="#4181a4", size=1.5) +
    labs(
        subtitle = "Randomly selected 80 genes w/o KRAB-ZNFs & 992 TEs",
        tag = "Figure",
        x = "percentage",
        y = "Count/Density",
    ) +
    annotate(
        geom = "text",
        x = 0.525,
        y = 32,
        label = "KRAB-ZNFs \nper. 52% \ndensity 0.89",
        hjust = 0
    ) +
    theme_minimal(
        base_size = 20
    )
```

#### compare all genes and all tes
```{r}
# create gene input data
df_gene <- gene
rownames(df_gene) <- gene$gene
df_gene <- df_gene[,c(2,3,4)]

# create te input data
df_te <- te
rownames(df_te) <- te$name
df_te <- df_te[,c(4,5,6)]

# run correlation, first 10 rows
corrOrthologTE(
    geneInput = df_gene[c(1:10), ],
    teInput = df_te[c(1:10), ],
    padjMethod = "fdr",
    fileDir = "hm_results"
)

```

