# TE expression investigation

In this part, we want to see if the expression of TEs that are negative correlated to 
KRAB-ZNFs actually lower than positive ones. First we import the data we need: expression data, correlation data.

```{r, warning=FALSE, message=FALSE}
# load library
library(twice)
library(dplyr)
```

```{r}
# expression data
hm_TE <- read.csv("data/counts/hm_te.csv")

# correlation data
hmCorr <- read.csv("../hm_results/hmCorrResult81.csv")
hmCorr.sig <- hmCorr %>%
    filter(padj < 0.05)

# filter TE name with positive or negative correlation
te_pos <- hmCorr.sig[hmCorr.sig$coef>0, ] #117 rows of data
te_neg <- hmCorr.sig[hmCorr.sig$coef<0, ] #178 rows of data

# unique number of TE
length(unique(te_pos$teName)) #86
length(unique(te_neg$teName)) #76
```
We have 117 KRAB-ZNFs-TEs significant positive correlations and 178 KRAB-ZNFs-TEs significant negative correlations from our correlation result obtained from TEKRABber.
From the filtering result, there are 86 TEs found in positive group and 76 TEs found in negative group.
